---
title: "p8105_hw5_ilz2105"
author: "Lulu Zhang"
date: 2019-09-17 
output: github_document
---


```{r setup, include = FALSE}
library(tidyverse)
```

Creating a data frame that has:
  a random sample of size 8 from a standard Normal distribution
  a logical vector indicating whether elements of the sample are greater than 0
  a character vector of length 8
  a factor vector of length 8, with 3 different factor “levels”

```{r df_create}

df_1 = tibble(
  ran_samp = rnorm(8),
  log_vec = ran_samp > 0,
  vec_char = c("I", "have", "no", "idea", "what", "Im", "doing" , "yikes"),
  fac_vec = factor(c( "yes", "no", "maybe", "yes", "no", "maybe", "yes", "no"))
)

view(df_1)

### Take the mean of each variable in dataframe

mean_samp = mean(pull(df_1, ran_samp))
mean_log = mean(pull(df_1, log_vec))
mean_char = mean(pull(df_1, vec_char))
mean_fac = mean(pull(df_1, fac_vec))


mean_samp
mean_log
mean_char
mean_fac

### Found the mean for `mean_samp` and `mean_log`
### Could not get the mean for `mean_char` and `mean_fac` because
### they were not numeric values 


```


Applying 'as.numeric' to the logical, character, and 
factor vectors to coerce into numeric vectors

```{r, eval = FALSE}

as.numeric(pull(df_1, vec_char))
as.numeric(pull(df_1, fac_vec))

### convert the logical vector to numeric, and multiply the random sample by the result

as.numeric(pull(df_1, log_vec))*(pull(df_1, ran_samp))

### convert the logical vector to a factor, and multiply the random sample by the result

as.factor(pull(df_1, log_vec))*(pull(df_1, ran_samp))

### This did not work because we cannot multiply a factor

### convert the logical vector to a factor and then convert the result to numeric, 
### and multiply the random sample by the result

as.numeric(as.factor(pull(df_1, log_vec)))*(pull(df_1, ran_samp))
```

--
Problem 2
--

Create a data frame comprised of:
x: a random sample of size 500 from a standard Normal distribution
y: a random sample of size 500 from a standard Normal distribution
A logical vector indicating whether x + y > 1
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector

### Create dataframe

```{r}

df_2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logical_vec = x + y > 1,
  num_log= as.numeric(logical_vec),
  fac_vec= as.factor(logical_vec)
)

view(df_2)
(mean(pull(df_2, x))*(pull(df_2, logical_vec))
  
(median(pull(df_2, x))*(pull(df_2, logical_vec))
  
(st(pull(df_2, x))*(pull(df_2, logical_vec))
```


`df_2` has `nrow(df_2)` rows and `5 columns`ncol(df_2)` columns
the mean of  `m`  
 

