---
title: "p8105_hw5_ilz2105"
author: "Lulu Zhang"
date: 2019-09-17 
output: github_document
---

--
__Problem 1__
--


```{r setup, include = FALSE}
library(tidyverse)
```

Created a data frame that has:
  a random sample of size 8 from a standard Normal distribution
  a logical vector indicating whether elements of the sample are greater than 0
  a character vector of length 8
  a factor vector of length 8, with 3 different factor “levels”

```{r df_create}

df_1 = tibble(
  ran_samp = rnorm(8),
  log_vec = ran_samp > 0,
  vec_char = c("I", "have", "no", "idea", "what", "Im", "doing" , "yikes"),
  fac_vec = factor(c( "yes", "no", "maybe", "yes", "no", "maybe", "yes", "no"))
)

### Take the mean of each variable in dataframe

mean_samp = mean(pull(df_1, ran_samp))
mean_log = mean(pull(df_1, log_vec))
mean_char = mean(pull(df_1, vec_char))
mean_fac = mean(pull(df_1, fac_vec))
```

Found the mean for `mean_samp` and `mean_log`, but could not get the mean
for `mean_char` and `mean_fac` because they were not numeric values.


Below, applied `as.numeric` to the logical, character, and 
factor vectors to coerce into numeric vectors.

```{r, eval = FALSE}

as.numeric(pull(df_1, vec_char))
as.numeric(pull(df_1, fac_vec))

### convert the logical vector to numeric, and multiply the random sample by the result
as.numeric(pull(df_1, log_vec))*(pull(df_1, ran_samp))

### convert the logical vector to a factor, and multiply the random sample by the result
as.factor(pull(df_1, log_vec))*(pull(df_1, ran_samp))

### convert the logical vector to a factor and then convert the result to numeric, 
### and multiply the random sample by the result

as.numeric(as.factor(pull(df_1, log_vec)))*(pull(df_1, ran_samp))
```
We can mutiply a factor so trying to convert `log_vec` to a factor and then multiplying
`ran_samp` by the result did not work. We are able to Convert `log_vec` to a factor and then convert 
that to a numeric vector and multiply that by `ran_samp` because they are both numeric vectors. The 
result was `r as.numeric(as.factor(pull(df_1, log_vec)))*(pull(df_1, ran_samp))`.




--
__Problem 2__
--

Created a data frame comprised of:
x, a random sample of size 500 from a standard Normal distribution,
y, a random sample of size 500 from a standard Normal distribution,
a logical vector indicating whether x + y > 1,
a numeric vector created by coercing the above logical vector,
and a factor vector created by coercing the above logical vector.


```{r}
df_2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logical_vec = x + y > 1,
  num_log= as.numeric(logical_vec),
  fac_vec= as.factor(logical_vec)
)
```


The dataframe created above, `df_2`, has `r nrow(df_2)` rows and `r ncol(df_2)` columns
The mean of `x` is `r mean(pull(df_2, x))`.
The median of `x` is `r mean(pull(df_2, x))`.
The standard deviation of `x` is `r sd(pull(df_2, x))`.
The proportion of cases for which x + y > 1 is `r mean(pull(df_2, logical_vec))`


Below I created a scatterplot of y vs x and colored points using the logical variable. 
Then, I created a second and third scatterplot that color points using the numeric and factor 
variables, respectively.

```{r yx_scatter}
# scatterplot with colored points using the logical variable
plot_1= ggplot(df_2, aes(x, y, color = logical_vec)) + geom_point()

# scatterplot with colored points using the numeric variable
plot_2 = ggplot(df_2, aes(x, y, color = num_log)) + geom_point()

# scatterplot with colored points using the factor variable
plot_3 = ggplot(df_2, aes(x, y, color = fac_vec)) + geom_point()


# export first scatterplot to directory
#ggsave('plot_1')
```

The color scheme was the same for the scatterplots that colored points using the logical variable `logical_vec`
and the factor variable `fac_vec` where it was two distinct colors, pink and blue to represent false and true. 
The color scheme for the scatterplot that colored points using the numeric variable `num_log` appeared to be 
more of a gradient from blue to black ranging from 0 to 1. However, on the plot, we only see blue or black 
which reprent 0 and 1 because we have no values in between because it was a dichotomous variable. 